# R Markdown {#rmarkdown}

First of all, I would like to say that markdown is amazing and that R markdown is even more amazing. 


## R markdown general tips for reports.

1. Firstly, when you load tidyverse you do not need to load the different packages contained in tidyverse such as dplyr, ggplot2, readr, tidytext separately. 

2. Secondly, all the libraries or packages should be in the same code chuck loaded in the beginning of the R markdown.

3. Everytime you successfully knit an R markdown file it will save the rendered out put as a separate file in either an html, word, or pdf file format. Hence, you can always open and see your last successfully rendered R markdown report. 



# Creating a new R Markdown File:
1. To create an R Markdown file, go to File -> New File -> R Markdown and hit enter.
2. Enter Title and Author of the document and press ok.
3. R Markdown file with the extension of .Rmd will appear with default code and comments, which can be removed.
4. Ctrl+s to save this file and name this R Markdown file accordingly.

# Commands used in R Markdown:
There are two places to write in R Markdown files:

#### i. Code blocks:
Text written in the highlighted part/ within \```{r} 
and \``` is the code.
#### ii. Comments blocks:
Remaining text is the comments.
Different paragraphs should be written with a line differences in the .Rmd file.
# Writing code in code blocks:
1. Press ctrl+alt+i to create a block to code in.
2. Code block will be created having \```{r} 
\```

#### i. To see both code and output on R Markdown, write code in between \```{r} and 
#### \``` 
For example:
The following code 
\``` {r} 
\ summary(cars) 
\``` 
will show both code and summary of table cars.
```{r}
summary(cars)
```

#### ii. If we only want to see the output and not the code then write the code between \```{r, 
#### \ echo=FALSE } 
#### and 
#### \```
For example:
The following code 

\```
\ {r , echo=FALSE}

\ summary(cars)

\```

will show the SUMMARY of table cars and not the code.

```{r, echo=FALSE}
summary(cars)
```


# Writing Comments:
* Anything that is written out of code block will be considered are comment and will be displayed on the screen.

# Types of Headings in R Markdown:
* Heading of size 1 can be achieved by writing # before the comment.
* For heading of size 2, two hash i.e: ## will be written before the comment and so on till heading 6. 
e.g: 

# This is written in Heading size 1. In html is h1.
## This is written in Heading size 2. In html is h2.
### This is written in Heading size 3. In html is h3.
#### This is written in Heading size 4. In html is h4.
##### This is written in Heading size 5. In html is h5.
###### This is written in Heading size 6. In html is h6.




## Sample paragraph displaying R Markdown capabilities

We describe our methods in this chapter.

Math can be added in body using usual syntax like this 

## math example

$p$ is unknown but expected to be around 1/3. Standard error will be approximated

$$
SE = \sqrt(\frac{p(1-p)}{n}) \approx \sqrt{\frac{1/3 (1 - 1/3)} {300}} = 0.027
$$

You can also use math in footnotes like this^[where we mention $p = \frac{a}{b}$].

We will approximate standard error to 0.027[^longnote]

[^longnote]: $p$ is unknown but expected to be around 1/3. Standard error will be approximated

    $$
    SE = \sqrt(\frac{p(1-p)}{n}) \approx \sqrt{\frac{1/3 (1 - 1/3)} {300}} = 0.027
    $$

