# SQL {#sql}


## SQL Overview

We will teach SQLite, Microsoft Sequal Server, and PostgreSQL.

### Basic SQL commands


### Selecting Columns and Dispalying the Data.

```{r, eval = F}

SELECT * FROM table;

SELECT column1, column2, column3, ... FROM table;

SELECT DISTINCT column_name FROM table;

```

### Counting the Rows and Groups.

```{r, eval = F}

SELECT COUNT(*) FROM table;

```


### Group by is Amazing.

```{r, eval = F}

SELECT COUNT(*) FROM table GROUP BY column1;

SELECT COUNT(*) FROM table GROUP BY column1, column2, ...;

```





Joining tables.

```{r, eval = F}

```




## Dr. Chuck Severance Questions

```{r, eval = F}

```


### Auto Increment

```{r, eval = F}

CREATE TABLE table (
  id SERIAL,
  name VARCHAR(128),
  email VARCHAR(128) UNIQUE,
  PRIMARY KEY(id)
);

```

### Time

Use the function NOW(). Also, use TIMESTAMP .

```{r, eval = F}

```


## PostgreSQL



Question 2
What is the psql command to show all of the databases in a server?
\l

What is the psql command to show the schema of a relation?
\dt  wrong

What is the psql command to show all the relations in a database?
\dd wrong



