--- 
title: "All About R and R Applications"
author: "Faizan Khalid Mohsin"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Prerequisites

Placeholder



<!--chapter:end:index.Rmd-->

# Introduction {#intro}


Hello Everyone! Hope you all are doing well. 

My hope is to help you become a data scientist via this book that you would otherwise need to learn to become by being hired and trained at a proper machine learning company. 


This book's examples and specific code is run on a window's operating system and is done using R. 






<!--chapter:end:01-intro.Rmd-->


# Overview

Placeholder



<!--chapter:end:02-overview.Rmd-->

# R Programming Techniques


## R Programming Essential Tips 

1. When repeating mutate to create for example percentage column in a table simply write that as a pip compatible function and call the function. 

2. Also, write small chucks of code and run it to see if it works. It will be much easier to debug if there is an error and to find and isolate it. It becomes very difficult to debug code the bigger it gets.


## Basics of R

### Print a string.
```{r, echo=TRUE, eval=TRUE}
print("First R lesson.")
```
### print a integer Value.

```{r}
print(23)
```

### Print with variable.

```{r}
x <- "First R lesson."
x
```

### Print String with varibale.

```{r}
x <- "First"
cat(x, "R lesson.")
sprintf("%s R lesson.",x)
```

## Assign variables

### Using = operator.

```{r}
x = 23
x
```

### Using <- operator.


```{r}
x <- "First program."
x
```

### Using -> operator.


```{r}
"First Program." -> x
x
```

### Multiple assignments

```{r}
x<- y<- 23
x
y
```


<!--chapter:end:03-rprogramming.Rmd-->


# Exploratory Data Analysis {#eda}

Placeholder


## Basic EDA Steps
### Step 1
### Step 2

<!--chapter:end:06-eda.Rmd-->

# General Useful Tips {#tips}

## Correct Writing


First of all, I want to start by Titles. It is very important that you know how titles are written properly. You need to title you analysis, visualization, applications, sections, everything correctly. 

Every first letter in a title should be capitalized except for prepositions or connecting words like and, or, the, this. 
If we want to convert the phrase general useful tips to a title it becomes: General Useful Tips. 

Now, also, sentences should be written properly. The first word's first letter should be capital and the sentence should end with a period also called full-stop like this one. 

## Essential Survival Habits.

One great thing I find about R Markdown and knitting a document is that to successfully knit there needs to be zero errors. So this is a great way to check if there is any error in the code.

For working in R Markdown. Please knit the markdown after making small changes in the code so you know if there is an error and can isolate, correct, and debug it. If you write a lot of code and then knit and there is a error, even a very simple error can be very hard to find buried in code. 

Further, once it knits successfully, then the code needs to be immediately committed to git and push it to GitHub.

These two steps of knitting and pushing the code to GitHub should be like breathing. Done extremely often automatically and even without thinking. If you can you should make a short cut like TAB + G so that the the r markdown knits and if successful then commits and pushes to GitHub only prompting you to add a commit message.

## Good General Habits.

Another, important tip is to look at your visualizations and the summary tables and think if the numbers you get make sense. I will be very frank here, I typically try not to do this because I have to use my brain and think. And sometimes that is actually very mentally power consuming.  

<!--chapter:end:07-tips.Rmd-->

# Git and Git Hub {#git}

First of all, I would like to say that command line and linux are amazing. It does things so simply and so easily it is almost as if you wave a hand a your room gets cleaned. You make a few hand gestures and wood, stone, and metal come together and assemble themselves into the frame of the house you have in mind. The ratio of the amount of effort you put in, the simplisity of the instructions and the accuracy of the execution to not what you say, but what you actually want is a kin to that of magic. So I write this ode in praise of the command line, Linux, and its predecessor Unix. 


## The Command Line

First please note that for the title the first letter of every word is capital except of connecting words which are not present. Also, since it is a Title it does not end with a peroid also called full-stop. 

Here are the basic commands to be used at the command line or terminal. 

## Essential Command Line Tips

Knowing these tips is essential because it increases productivity by so much and makes this much easier to do. 

1. **Tip Number 1:** When simultaneously working on two different projects, not even necessarily programming projects, then have two terminals or git bash terminals open each one cd'ed into the respepective project's working directory. 

I tend to work on two projects at the same time. This is what I found to be very productive for me. I know if I start becoming less productive i.e. start not having ideas, hit bottle necks, become very slow, then I know I can switch to the other project. Also, when I only have one project open I feel a lot more pressure, that as if I need to finish it or complete it quickly. By having two projects open I experience less stress and don't feel like if I am unable to progress in one then I am stuck. Have two projects open removes the feeling of only having one way forward and the fear of being stuck if that way is blocked. 

So when you have two separate coding projects running and you need to commit and push the changes to the GitHub, the amazing thing is to just have two git bash or terminals open with each being in one project and they run independently of one another. For some applications you cannot have two separate intinces running at the same time doing separate things. 

<!--chapter:end:08-git.Rmd-->


# R Markdown {#rmarkdown}

Placeholder


## R markdown general tips for reports.
## Sample paragraph displaying R Markdown capabilities
## math example

<!--chapter:end:09-rmarkdown.Rmd-->


# Visualization {#viz}

Placeholder


## Essential Visualization Principals
## Writing Visualization Titles
## Bar Chart tips

<!--chapter:end:10-visualization.Rmd-->

# Final Words

We have finished a nice book.

<!--chapter:end:100-summary.Rmd-->

# Applications

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:19-applications.Rmd-->

`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:20-references.Rmd-->

